\chapter{\label{ch:2-Methods section}Methods section}

\minitoc

\section{Cell culture procedures}

\subsection{Routine cell culture}

Huh7 cells \textbf{from where?} were cultured in maintenance media made from Dulbecco's modified Eagle's medium (DMEM) supplemented with GlutaMAX\textsuperscript{TM}, 5.5 mmol/L glucose, 1 mM sodium pyruvate, 2.4 g/L sodium bicarbonate (Gibco, Life Technologies), 10\% foetal bovine serum (FBS), 10,000 U/ml penicillin-streptomycin (P/S) and 1\% non-essential amino acids (NEAA). Maintenance media was changed every 2-3 days and cells kept in a incubator at 37\textdegree C and 5\% carbon dioxide. 

\subsection{Passaging cells}

Cells were grown to at least 80\% confluence in T175 (175cm\textsuperscript{2}) flasks before being washed with phosphate buffered saline (PBS) and then incubated for 5mins in 4mL of TrypLE\textsuperscript{TM}. Flasks were then agitated to dislodge cells and 10mL of maintenance media was added. Cells were then centrifuged for at 800rpm for 3mins at 21\textdegree C. The supernatant was removed and the pellet re-suspended in 5mL of maintenance media before being seeded 1:5 in new T175 flasks with maintenance media.

\subsection{Seeding cells for experiments}

Cells were isolated and re-suspended in maintenance media as described for passaging. 20\uL of cells were added to both sides of a cell counter slide and cells were counted using a Cellometer Auto T4 Bright Field Cell Counter. An average of both sides was taken and cells were then diluted with maintenance media to a concentration of 1 x 10\textsuperscript{5} cells per mL. Total cell number per well were as follows: 6-well plate - 2 x 10\textsuperscript{5}, 12-well plate - 1 x 10\textsuperscript{5}, 24-well plate - 0.5 x 10\textsuperscript{5}.

\subsection{Freezing cells for storage in liquid nitrogen}

\subsection{Thawing cells from storage in liquid nitrogen}

Frozen cells were placed in a 37\textdegree C water bath to thaw rapidly. 10mL of maintenance media was then added and cells were centrifuged at 800rpm for 3mins at 21\textdegree C. The supernatant was removed and the pellet re-suspended in 5mL of maintenance media and seeded 1:1 in a T75 (75cm\textsuperscript{2}) flask with maintenance media. Media was changed the next day and then every other day until cells reached at least 80\% and were passaged into a T175 flask.

\section {Quantitative reverse-transcription PCR}

\subsection{RNA extraction}

Cells for were cultured in 12-well plates and upon harvesting were washed twice with warm PBS before 250\uL of TRI Reagent\textsuperscript{\textregistered}  was added to each well. Cells were dislodged from the plate using a cell scraper and transferred into Eppendorf tubes to be frozen at -80\textdegree C until required. When required, samples were added to Phase Lock Gel tubes with 50\uL of chloroform and shaken vigorously. Samples were left for 5mins and then centrifuged for 15mins at 1200g and 4\textdegree C. The clear supernatant was transferred to 2mL tapered-bottom elution tubes containing 2\uL glycol blue and 200\uL isopropanol before being left overnight at -20\textdegree C. 

Samples were then centrifuged at 1200g and 4\textdegree C for 30mins to allow a RNA pellet to form. The supernatant was removed and pellet resuspended in 80\% ethanol before being centrifuged again at 1200g and 4\textdegree C for 15mins. The ethanol was then removed and the pellet allowed to dry before re-suspension in 15\uL RNAse free water and incubation for 3hours. RNA purity and concentration was then measured using a NanoDrop ND-1000 Spectrophotometer.

\subsection{Reverse transcription}

The volume of sample containing 1ug RNA was transferred into a 96 well plate and made up to 10uL with RNAse-free water. For each well with sample, 2\uL 10x reverse transcription buffer, 0.8\uL 25x deoxynucleoside triphosphates (dNTPs), 2\uL 10x random primers, 1\uL reverse transcriptase and 4.2\uL RNAse-free water was added. The plate was sealed and briefly spun before being placed on an MJ Research Tetrad PTC-225 Thermal Cycler. The thermal cycler heated the plate to 27\textdegree C for 10 minutes for the random primers to anneal to the sample RNA and this was then followed by 120 minutes at 37\textdegree C for the reverse transcriptase to form the cDNA using the dNTPs. Finally the reaction was terminated by heating the plate to 85\textdegree C for 5 minutes. The cDNA could then be stored in the freezer until required. 

\subsection{Quantative PCR}

The samples were diluted 1:40 in 10 mmol/L Trizma\R hydrochloride and measured in triplicate for each gene. A standard curve was made through serial dilution of cDNA pooled from all of the samples in 10 mmol/L Trizma\R and included for each gene. Quantative PCR was carried out on \textbf{Insert name of machine} and TaqMan\R assays used to measure gene expression. A 6\uL reaction volume was used consisting of 2.7\uL of sample cDNA, 0.3\uL TaqMan\R probes and 3\uL 2X KAPA PROBE FAST Master Mix. Relative expression was calculated as a ratio from generated Ct values as previously described \cite{Pfaffl2001ARTPCR}. Each sample was expressed relative to a calibrator sample and corrected to the geometric mean of three reference housekeeper genes (B2M, HMBS, and YWHAZ).